---
title: "Lab 15"
author: "Drew Hoffman, David Washburn, Tanner Ball, Brandon Nelsen, Burke Parent"
date: "12/10/2019"
output: html_document
---

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(modelr)
library(ggrepel)
```


```{r}
stats <- read.csv(file="bballdata.csv", header=TRUE, sep=",")
stats$Player <- sub("\\\\.*", "", stats$Player)
```

***Overall Question***
What is the most important statistic to get playing time for each position?

*Why It's Important*
This question is important because it can help us determine what statistics are necessary for a given player's position to improve their game.


**Conclusion**
git good lol

**Recommendation**
dont be bad

*Graph*


***Individual Sections***

*Tanner:*

Question: How closely do teams follow the "Four Factors" idea when determining players' minutes? The Four Factors come from one of the original NBA analytics guys, Dean Oliver. He claimed that the Four Factors are a philosophy that correlates strongly to a team's success in a game. The Four Factors are as follows: (1) Score every possession, (2) Pick up offensive rebounds, (3) Get to the free throw line, and (4) Protect the ball. 

These are philosophies that most coaches have been following for a long time, but I'm curious to see if it actually translates to playing time for their players. We will use effective FG% (eFG.), which is an adaptation of normal FG% that values three-pointers higher, Offensive Rebounds (ORB), free throw attempts (FTA), and turnovers (TOV), as well as minutes played per game (MP/G) to explore this question.

```{r}
four_factors <- stats %>% select(Player, Tm, MP, G, eFG., ORB, FTA, TOV)
# initial plots to look for interesting trends
ggplot(four_factors, aes(eFG., MP/G, color = Tm)) + geom_line()
ggplot(four_factors, aes(ORB, MP/G, color = Tm)) + geom_line()
ggplot(four_factors, aes(FTA, MP/G, color = Tm)) + geom_line()
ggplot(four_factors, aes(TOV, MP/G, color = Tm)) + geom_line()

all_mod <- lm(MP/G ~ eFG. + ORB + FTA + TOV, data = four_factors)
summary(all_mod)
```

*David*

Russell Westbrook has been called one of the most overrated players in the NBA, but do his numbers support this claim? If the data support the conclusion that he is an average player, we will then claim he is overrated. Hence, our null hypothesis will be: Russel Westbrook is an average or worse NBA player. Our Alternate will be: Russel Westbrook is better than the average NBA player. In order to conduct this experiment, the stats I will be looking at will be points per game and minutes played per game.His points can tell us how good he is offensively and his minutes can give us an idea of how much respoonsibility the coaching staff gives him.

Question: Is Russel Westbrook's "Overrated" status desserved? This question is interesting because we can use statistics to explain whether or not a specific player is as good as the average viewer thinks they are.

New Tool: Permutation Test

Answer: Based on the histograms and data below, we reject the null hypothesis in favor of the alternate. In essence, Russel Westbrook is not overrated and his points per game and minutes played per game support this claim. This relates to the overall question in that we are doing a study of a specific player's position. Since we have data that support Russel Westbrook is good at basketball, we can use him as an example for what other point guards should do to help them improve their game.

Ethical Implications: The ethical implication of these findings are that I am using numbers to judge a person. This probably is not ethically right as there could certainly be other factors that contribute to how well somebody plays a certain position.

```{r}

#make a points per game and minutes per game col

stats2 <- mutate(stats, ppg = PTS/G) %>%
          mutate(mppg = MP/G)
  

#Compare his stats to other players of the same position

Westbrook <- filter(stats2, Rk == 506)
Pointgaurds <- filter(stats2, Pos == 'PG', G > 29) #ppg = 22.95, mppg = 36.03

#get some averages

mean(Pointgaurds$ppg) # 10.85
mean(Pointgaurds$mppg) # 23.51

#plot of pointgaurds' ppg and mppg

ggplot() +
  geom_point(Pointgaurds, mapping = aes(mppg, ppg), color = "blue") +
  geom_point(Westbrook, mapping = aes(mppg, ppg), color = "red") +
  ggtitle("NBA Pointguards' points per game vs minutes played per game") +
  xlab("Minutes Played per Game") +
  ylab("Points Per Game")

#do a permuation test by sampling a bunch of ppg and mppg statistics for randomly sampled groups

  #ppg test

ppg <- select(Pointgaurds, ppg)

vector.ppg <- vector("double", length = 100)
for(i in seq_along(vector.ppg)){
  vector.ppg[i] <- sample_n(ppg, 1)
}


ppg.df <- unlist(vector.ppg)
  
  #mppg test 

mppg <- select(Pointgaurds, mppg)

vector.mppg <- vector("double", length = 100)
for(i in seq_along(vector.mppg)){
  vector.mppg[i] <- sample_n(mppg, 1)
}

mppg.df <- unlist(vector.mppg)

#ppg histogram
ggplot() + geom_histogram(aes(ppg.df), bins = 20) +
    geom_vline(xintercept =  mean(ppg.df), color = "purple") + #mean of sampled data
    geom_vline(xintercept = 23, color = "red") + #Westbrook 
    ggtitle("Distribution of NBA Points Per Game among Pointguards") +
    xlab("Points per Game")

#mppg histogram
ggplot() + geom_histogram(aes(mppg.df), bins = 20) +
    geom_vline(xintercept =  mean(mppg.df), color = "purple") + #mean of sampled data
    geom_vline(xintercept = 36, color = "red") + #Westbrook 
    ggtitle("Distribution of NBA Minutes Played Per Game among Pointguards") +
    xlab("Minutes played per Game")

```

**Team Reflection**

**Individual Reflections**

*Davd*
At the start of this course my goal for 6 months after college was to be travelling in Europe, my 5 year goal was that I wanted to be analayzing statistics. These goals haven't changed for me, but I do feel that I am more commited to achieving my 5 year goal as a result of this course. I achieved at lease a basic understanding of R in this course, considering that this is my first experience with data science and coding, I'm pretty satisfied with this result. The advice that I would give myself at the beginning of this semester would be to get a good understanding of the examples in the book.
