---
title: "Lab 15"
author: "Drew Hoffman, David Washburn, Tanner Ball, Brandon Nelsen, Burke Parent"
date: "12/10/2019"
output: html_document
---

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(modelr)
library(stringr)
library(ggrepel)
```


```{r}
stats <- read.csv(file="bballdata.csv", header=TRUE, sep=",")
stats$Player <- sub("\\\\.*", "", stats$Player) # keeps only the player's full name
```

**Overall Question**
What is the most important statistic to get playing time for each position?

**Why It's Important**
This question is important because it can help us determine what statistics are necessary for a given player's position to improve their game.


**Conclusion**

For guards, we can conclude that they need more than just strong shooting to get playing time. In fact, guards don't necessarily have to be good shooters at all to get time, as long as they are strong in other areas, such as assists. We see this case highlighted by players like Ben Simmons, an All-Star who didn't hit a three-pointer all of last season, and Russell Westbrook, a below-average three-point shooter who often is called overrated, but still gets significant time because of his superior all-around game.

**Recommendation**
dont be bad

```{r}
PTVariable <- stats %>% filter(MP > 100) %>% select(Age,MP,FG.,X3P.,X2P.,FT.,STL,BLK,PTS)

cor(stats$MP, stats$Age)
cor(stats$MP, stats$FG)
cor(stats$MP, stats$X3P)
cor(stats$MP, stats$X2P)
cor(stats$MP, stats$FT)
cor(stats$MP, stats$STL)
cor(stats$MP, stats$BLK)
cor(stats$MP, stats$PTS)

ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = Age))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = FG.))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = X3P.))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = X2P.))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = FT.))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = STL))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = BLK))
ggplot(data = PTVariable)+
  geom_point(aes(x = MP, y = PTS))
```



# Individual Sections:

### **Tanner:**

**Question:** How critical is it for guards to be able to shoot the ball in order to get playing time?

In an era where players are shooting more than ever, how much of a factor is shooting for guards to get playing time? Guards are typically seen as the better shooters on their team, so I was curious to see if playing time will always go up as shooting percentage goes up. For our dataset, we will be looking at guards who played at least 18 minutes per game, as these guys can be considered rotation players who regularly get in the game.

We will be looking at three types of field goal percentages: (1) Effective Field Goal Percentage (eFG%), an adaptation of normal field goal percentage, which slightly favors three point shooting, (2) 2-Point Field Goal Percentage (2FG%), or the shooting percentage of our players on 2-point shots, and (3) 3-Point Field Goal Percentage (3FG%), which is the shooting percentage of our players on 3-point shots. 

**New Tools:**  Linear Modeling

We will also be pointing out three players in particular to see where they stand in our three models. Our three players are (1) Ben Simmons of the Philadelphia 76ers, an All-Star notorious for almost never shooting from outside the paint, (2) Stephen Curry of the Golden State Warriors, an All-Star who is arguably the greatest shooter of all time, and (3) Kyle Korver of the Cleveland Cavaliers and Utah Jazz, an aging veteran who is pretty much a 3-point specialist. 

**Answer:** What we see isn't particularly surprising. Our predictions indicate that the better you shoot, the more minutes per game you are likely to play. However, when we look at our three highlighted players, we see something a little more interesting. 

Ben Simmons, despite not hitting a single three-pointer all of last season, finished with a higher eFG% than sharpshooter Kyle Korver. This could be because Simmons played far more minutes than Korver, but if we look at two-point shooting we see that Simmons was far better than Korver inside the arc. In our three-point shooting chart, we can see that Simmons' inepitude from distance had seemingly no effect on his minutes per game. This outlier surely had a large effect on our prediction line for three-point shooting as well, bringing its starting point up much higher than it might be otherwise. There are also a few other points we can see that indicate that below-average three-point shooting does not necessarily hurt playing time.

Kyle Korver, despite being an outstanding three-point shooter, does not get much playing time at all. At this stage of his career, it seems he is a middling role player at best. Possible reasons for this could be his age -- at 37 years old, he is among the oldest active players in the league -- or perhaps he falls far short in other areas of his game, hurting his ability to get playing time.

This makes the case of Stephen Curry pretty interesting. Curry is without a doubt an incredible three-point shooter, finishing last season 6th in 3FG% among guards and with an eFG% greater than 60%, which is astounding given his playing time. However, as we just saw with Korver, there must be other factors at play to determine playing time. It may be fair to say that although Curry's red-hot shooting certainly gets him a lot of minutes, he must also be strong in other aspects of his game to warrant so much more playing time than most of his peers at his position.

Based on our models and our look at these particular players, I think we can say that for most players, stronger shooting will get them more time on the floor (and in the case of players like Korver, it may be the only reason they're out there). However, guards with strong ability in other aspects of their game can likely get by without great shooting. Ben Simmons did not hit a single three-pointer all season and still finished among the highest in minutes played and was an NBA All-Star for the 76ers last season. Shooting is not the be-all and end-all for guards.

**Ethical Implications:** I think something we can take away from these findings is that coaches, players, executives, and fans should not be so quick to judge a player solely on his ability to shoot the ball. Shooting has certainly grown in importance in the minds of fans and organizations, so it's understandable how this happens, but we should all still try to look at a player's full set of skills before determining their "rank" among their peers. Although a player should certainly work on his shooting, there are other ways to be a great player while the shot is still a work-in-progress.


```{r}
# Set up our dataset - We will be looking at guards who played at least 18 mins/g (rotation players)
guards <- stats %>% filter(str_detect(Pos, "PG|SG"))
guards <- guards %>% mutate("MPG" = MP/G, "PPG" = PTS/G)
guards <- guards %>% filter(MPG >= 18)

# Three guards I wanted to look at in particular
Simmons <- guards %>% filter(Player == "Ben Simmons") # All-Star who never shoots 3s
Curry <- guards %>% filter(Player == "Stephen Curry") # All-Star who is arguably the greatest shooter ever
Korver <- guards %>% filter(Player == "Kyle Korver") # 37-year-old 3pt specialist

# Set up our models
efg_mod <- lm(MPG ~ eFG., data = guards)
twos_mod <- lm(MPG ~ X2P., data = guards)
threes_mod <- lm(MPG ~ X3P., data = guards)

# Setting up predictions
efg_grid <- guards %>% data_grid(eFG.)
efg_grid <- efg_grid %>% add_predictions(efg_mod, var = "eFG_pred")
two_grid <- guards %>% data_grid(X2P.)
two_grid <- two_grid %>% add_predictions(twos_mod, var = "two_pred")
three_grid <- guards %>% data_grid(X3P.)
three_grid <- three_grid %>% add_predictions(threes_mod, var = "three_pred")

# eFG% by MPG
ggplot(guards, aes(eFG.)) + geom_point(aes(y = MPG)) + geom_line(aes(y=eFG_pred), data=efg_grid, color="red", size=1) + geom_point(data=Simmons, aes(y=MPG), color = "blue") + geom_point(data=Curry, aes(y=MPG), color = "goldenrod1") + geom_point(data=Korver, aes(y=MPG), color = "purple2") + geom_text_repel(data=guards, aes(y=MPG, label=ifelse(Player == "Stephen Curry" | Player == "Ben Simmons" | Player ==  "Kyle Korver" ,as.character(Player),'')), force=5) + ggtitle("Minutes per Game by Effective Field Goal Percentage", "2018-19 Point Guards and Shooting Guards") + xlab("Effective Field Goal Percentage") + ylab("Minutes per Game")

# 2pt FG% by MPG
ggplot(guards, aes(X2P.)) + geom_point(aes(y = MPG)) + geom_line(aes(y=two_pred), data=two_grid, color="red", size=1) + geom_point(data=Simmons, aes(y=MPG), color = "blue") + geom_point(data=Curry, aes(y=MPG), color = "goldenrod1") + geom_point(data=Korver, aes(y=MPG), color = "purple2") + geom_text_repel(data=guards, aes(y=MPG, label=ifelse(Player == "Stephen Curry" | Player == "Ben Simmons" | Player ==  "Kyle Korver" ,as.character(Player),'')), force=10) + ggtitle("Minutes per Game by 2-Point Field Goal Percentage", "2018-19 Point Guards and Shooting Guards") + xlab("2-Point Field Goal Percentage") + ylab("Minutes per Game")

# 3pt FG% by MPG
ggplot(guards, aes(X3P.)) + geom_point(aes(y = MPG)) + geom_line(aes(y=three_pred), data=three_grid, color="red", size=1) + geom_point(data=Simmons, aes(y=MPG), color = "blue") + geom_point(data=Curry, aes(y=MPG), color = "goldenrod1") + geom_point(data=Korver, aes(y=MPG), color = "purple2") + geom_text_repel(data=guards, aes(y=MPG, label=ifelse(Player == "Stephen Curry" | Player == "Ben Simmons" | Player ==  "Kyle Korver" ,as.character(Player),'')), force=5) + ggtitle("Minutes per Game by 3-Point Field Goal Percentage", "2018-19 Point Guards and Shooting Guards") + xlab("3-Point Field Goal Percentage") + ylab("Minutes per Game")
```

### **David:**

Russell Westbrook has been called one of the most overrated players in the NBA, but do his numbers support this claim? If the data support the conclusion that he is an average player, we will then claim he is overrated. Hence, our null hypothesis will be: Russel Westbrook is an average or worse NBA player. Our Alternate will be: Russel Westbrook is better than the average NBA player. In order to conduct this experiment, the stats I will be looking at will be points per game and minutes played per game.His points can tell us how good he is offensively and his minutes can give us an idea of how much respoonsibility the coaching staff gives him.

**Question:** Is Russel Westbrook's "Overrated" status desserved? This question is interesting because we can use statistics to explain whether or not a specific player is as good as the average viewer thinks they are.

**New Tool:** Permutation Test

**Answer:** Based on the histograms and data below, we reject the null hypothesis in favor of the alternate. In essence, Russel Westbrook is not overrated and his points per game and minutes played per game support this claim. This relates to the overall question in that we are doing a study of a specific player's position. Since we have data that support Russel Westbrook is good at basketball, we can use him as an example for what other point guards should do to help them improve their game.

**Ethical Implications:** The ethical implication of these findings are that I am using numbers to judge a person. This probably is not ethically right as there could certainly be other factors that contribute to how well somebody plays a certain position.

```{r}

#make a points per game and minutes per game col

stats2 <- mutate(stats, ppg = PTS/G) %>%
          mutate(mppg = MP/G)
  

#Compare his stats to other players of the same position

Westbrook <- filter(stats2, Rk == 506)
Pointgaurds <- filter(stats2, Pos == 'PG', G > 29) #ppg = 22.95, mppg = 36.03

#get some averages

mean(Pointgaurds$ppg) # 10.85
mean(Pointgaurds$mppg) # 23.51

#plot of pointgaurds' ppg and mppg

ggplot() +
  geom_point(Pointgaurds, mapping = aes(mppg, ppg), color = "blue") +
  geom_point(Westbrook, mapping = aes(mppg, ppg), color = "red") +
  ggtitle("NBA Pointguards' points per game vs minutes played per game") +
  xlab("Minutes Played per Game") +
  ylab("Points Per Game")

#do a permuation test by sampling a bunch of ppg and mppg statistics for randomly sampled groups

  #ppg test

ppg <- select(Pointgaurds, ppg)

vector.ppg <- vector("double", length = 100)
for(i in seq_along(vector.ppg)){
  vector.ppg[i] <- sample_n(ppg, 1)
}


ppg.df <- unlist(vector.ppg)
  
  #mppg test 

mppg <- select(Pointgaurds, mppg)

vector.mppg <- vector("double", length = 100)
for(i in seq_along(vector.mppg)){
  vector.mppg[i] <- sample_n(mppg, 1)
}

mppg.df <- unlist(vector.mppg)

#ppg histogram
ggplot() + geom_histogram(aes(ppg.df), bins = 20) +
    geom_vline(xintercept =  mean(ppg.df), color = "purple") + #mean of sampled data
    geom_vline(xintercept = 23, color = "red") + #Westbrook 
    ggtitle("Distribution of NBA Points Per Game among Pointguards") +
    xlab("Points per Game")

#mppg histogram
ggplot() + geom_histogram(aes(mppg.df), bins = 20) +
    geom_vline(xintercept =  mean(mppg.df), color = "purple") + #mean of sampled data
    geom_vline(xintercept = 36, color = "red") + #Westbrook 
    ggtitle("Distribution of NBA Minutes Played Per Game among Pointguards") +
    xlab("Minutes played per Game")

```

### **Burke:**

**Question:** How do assists per player attribute to total playing time?

**New Tool:** Data Transformation and Pipes

**Answer:** After modeling the data using a dot plot, their appears to be a trend in which higher number of assists is positively correlated with a higher number of minutes played total. This relationship, however, is likely not causal, because more time played on the court obviously gives players more opportunities to rack up assists.

**Ethical Implications:** This data is fairly obvious, and I do not see how it could be abused in any way, such as influencing future NBA games or drafts. Athletes are always judged based on their statistics over their career, and businesses capitalize on our obsession with sports through live betting, fantasy drafts, and more. Looking at a players statistics and noticing overall trends is a practice that has been exercised for years, and I would say that if it ever did have an effect on athletes, they get used to this scrutinization of their playing as their career progresses.

```{r}
AST_MP <- stats %>% select(Player, Tm, MP, AST,)
AST_MP

#Plot Detailing Trend#
ggplot(AST_MP, mapping = aes(MP, AST))+
  geom_point()+geom_smooth(se = FALSE)+
  xlab("Minutes Played per Player")+
  ylab("Assists per Player")

#Looking at the five players with the most assists.#
High_AST <- AST_MP %>% arrange(desc(AST)) %>% mutate(top5 = AST > 585)
#Removing other Players from Data Frame.#
High_AST <- High_AST %>% filter(top5 == 1)

ggplot(High_AST, mapping = aes(MP, AST, color = Player))+geom_point()

```

### **Drew:**

**Question:** How important is it for an NBA center to block shots and grab defensive rebounds while they are on the court?

The NBA center is usually the backbone of any defensive unit. While offenses utilize their screen setting ability and size, holding their own on the defensive end is key for any NBA big man. In fact, 25 of the 37 Defensive Player of the Year awards have been awarded to a player that plays significant time at the position. Their role as an anchor and a big body that patrolls the paint allows the center to disrupt shots at the rim and make it harder for the other team to score close to the basket. They also need to be able to secure defensive rebounds in order to end opposing team's possessions and get the ball in the hands of their teammates. This begs the question, how proficient does a center need to be at blocking shots and grabbing boards to get time on the court?

To answer this, I will be normalizing the blocks and defensive rebounding stats of centers for a 36 minute pace and then regressing it against their minutes played per game. From this, we will see if there is a significant correlation between the variables. From there, we can determine if blocking shots or getting rebounds are skills that an NBA center needs in order to get time on the court.

However, this will not tell the whole story. Centers in the NBA play traditionally fewer minutes than the average player due to the nature of their physical play and conditioning. It is easier for a smaller guard or quick wing player to stay on the court when they aren't constantly battling for post position, setting screens, or boxing out other 7 footers for rebounds. Often centers will start the game to set the defensive tone and create favorable matchups but not play a large quantity in the end. Because of this, I will also be regressing the blocks and rebounds data against "Start rate", or the proportion of games that the center started out of all games played. This will help give a holistic view of their value to the team that minutes per game might hide.

Another caveat is that defensive rebounds aren't telling the whole story either. Players like Steven Adams are known for being very adept at boxing out players and allowing their teammates to secure the ball and push it up court faster than they would be able to. To improve this analysis, I would also need box out data, but this is a new statistic and is not standardized across the league.

Null hypotheses: 

There is no correlation between block rate and minutes played per game.

There is no correlation between defensive rebounding rate and minutes played per game.

Alternate hypotheses:

There is a correlation between block rate and minutes played per game.

There is a correlation between defensive rebounding rate and minutes played per game.

**New Tool:** Correlation testing

**Answer:** From the analysis, we can see that while block rate is not statistically significant in a player receving playing time, their rate of grabbing defensive rebounds is. The regression tests for block rate fall comfortably within the distribution so we fail to reject the null hypothesis that block rate has no effect on playing time. However, the regression test for defensive rebounding rate shows a statistically significant positive correlation between grabbing the rebound to end the possession and staying out on the court.

**Ethical Implications:** This information shows how difficult it is to quantify a player's defensive value on the court. Blocks seem like they would do a good job of showing who is valuable on defense and who isn't, but as we can see blocking shots doesn't necessarily mean that a player is going to play more. Defensive rebounds provide more tangible value, but as I explained earlier there is more to the story. In order to improve player evaluation tools, we need more comprehensive statistics compared with team success as well as a holistic knowledge of the game to find out what the stats are telling. People tend to twist numbers in a way that supports their own narratives and opinions, in a way that extends past sports and into the real world. Recognizing that helps disseminate claims that are poorly supported. As with anything, understanding context is vital and there is a need for intellectual honesty when sharing data.

```{r}
#filter based on significant playing time and position
centers <- filter(stats, Pos == 'C', MP > 200)

#normalize blocks for 36 minute pace and caculate MPG
centers <- mutate(centers, brate = (BLK/MP)*36) %>%
  mutate(mppg = MP/G) %>%
  mutate(srate = GS/G) %>%
  mutate(drbrate = (DRB/MP)*36)

#extracting vectors for testing
blockrates <- centers$brate
minutespg <- centers$mppg
startrate <- centers$srate
defrebrate <- centers$drbrate

#correlation permutation code
perm_cor <- function(perms = 1000, x, y)
{
  cor_coeffs <- numeric(perms)

  for (i in c(1:perms))
  {
    mix_up <- sample(y)

    coefficient <- cor(x, mix_up)
    cor_coeffs[i] <- coefficient
  }
  
  return(cor_coeffs)
}

#correlation test
blocksmpgfakes <- perm_cor(1000, blockrates, minutespg)
blockssratefakes <- perm_cor(1000, blockrates, startrate)
blocksmpgreal <- cor(blockrates, minutespg)
blockssratereal <- cor(blockrates, startrate)

ggplot() + geom_histogram(aes(blocksmpgfakes)) +
  geom_vline(xintercept = blocksmpgreal) +
  ggtitle("Correlation Permutation Test Between Block Rate and Minutes Per Game") +
  xlab("Correlation") +
  ylab("Distribution")

ggplot() + geom_histogram(aes(blockssratefakes)) +
  geom_vline(xintercept = blockssratereal) +
  ggtitle("Correlation Permutation Test Between Block Rate and Start Rate") +
  xlab("Correlation") +
  ylab("Distribution")

boardsmpgfakes <- perm_cor(1000, defrebrate, minutespg)
boardssratefakes <- perm_cor(1000, defrebrate, startrate)
boardsmpgreal <- cor(defrebrate, minutespg)
boardssratereal <- cor(defrebrate, startrate)

ggplot() + geom_histogram(aes(boardsmpgfakes)) +
  geom_vline(xintercept = boardsmpgreal) +
  ggtitle("Correlation Permutation Test Between Defensive Rebounding Rate and Minutes Per Game") +
  xlab("Correlation") +
  ylab("Distribution")

ggplot() + geom_histogram(aes(boardssratefakes)) +
  geom_vline(xintercept = boardssratereal) +
  ggtitle("Correlation Permutation Test Between Defensive Rebounding Rate and Start Rate") +
  xlab("Correlation") +
  ylab("Distribution")

```

### **Brandon**

**Question:** What shooting are most affected by a players age?

**New Tool:** QPlot and grid.arrange (Data Manipulation)

**Answer:** Looking at the graphs and the means it is clear that a player in their Late 20s, 25 to 30, has the best mean stats for any shooting statistic. I kept the NA in the above 35 category in the 3 point shooting because I thought it would make sense that many older players don't take three point shots, this could be due to their shooting skills declining after being in the league for so long. The young players both under 20s and early 20s are very close to the late 20s players, but did not surpass them in any area. The only area that the late 20s were passed in was the mean of the late 30s players in three points, this could possibly be that the slightly older players like Rajon Rondo who is a three point specialist, just play their single position and shoot a lot more or are possibly more accurate than the late 20s in the three point shooting area.

**Ethical Implications:** This could be useful for coaches and teams looking to pick up older players because it is always a risk to choose someone older because they are more prone to injury and they may start declining after a few years especially in their shooting overall, it would be much more beneficial to pick up a new young player and have them train and learn to become a better player so they can achieve their peak preformance in their late 20s.

```{r}
Under20 <- stats %>%  filter(Age < 20) %>% select(Age,FG,FGA,FG.,X3P,X3PA,X3P.,X2P,X2PA,X2P.,FT,FTA,FT.)
Early20s <- stats %>% filter(Age >= 20 & Age < 25)%>% select(Age,FG,FGA,FG.,X3P,X3PA,X3P.,X2P,X2PA,X2P.,FT,FTA,FT.)
Late20s <- stats %>% filter(Age >= 25 & Age < 30)%>% select(Age,FG,FGA,FG.,X3P,X3PA,X3P.,X2P,X2PA,X2P.,FT,FTA,FT.)
Early30s <- stats %>% filter(Age >= 30 & Age < 35)%>% select(Age,FG,FGA,FG.,X3P,X3PA,X3P.,X2P,X2PA,X2P.,FT,FTA,FT.)
Above35 <- stats %>% filter(Age >= 35)%>% select(Age,FG,FGA,FG.,X3P,X3PA,X3P.,X2P,X2PA,X2P.,FT,FTA,FT.)

FGDF <- data.frame("Age" = c("Under 20", "Early 20s", "Late 20s","Early 30s","Above 35"), "Field Goals" = c(mean(Under20$FG), mean(Early20s$FG), mean(Late20s$FG), mean(Early30s$FG), mean(Above35$FG)) , "Field Goal Attempts" = c(mean(Under20$FGA), mean(Early20s$FGA), mean(Late20s$FGA), mean(Early30s$FGA), mean(Above35$FGA)))

P3DF <- data.frame("Age" = c("Under 20", "Early 20s", "Late 20s","Early 30s","Above 35"), "Three Points" = c(mean(Under20$X3P), mean(Early20s$X3P), mean(Late20s$X3P), mean(Early30s$X3P), mean(Above35$FX3P)) , "Three Point Attempts" = c(mean(Under20$X3PA), mean(Early20s$X3PA), mean(Late20s$X3PA), mean(Early30s$X3PA), mean(Above35$X3PA)))

P2DF <- data.frame("Age" = c("Under 20", "Early 20s", "Late 20s","Early 30s","Above 35"), "Two Points" = c(mean(Under20$X2P), mean(Early20s$X2P), mean(Late20s$X2P), mean(Early30s$X2P), mean(Above35$X2P)) , "Two Point Attempts" = c(mean(Under20$X2PA), mean(Early20s$X2PA), mean(Late20s$X2PA), mean(Early30s$X2PA), mean(Above35$X2PA)))

FTDF <- data.frame("Age" = c("Under 20", "Early 20s", "Late 20s","Early 30s","Above 35"), "Free Throws" = c(mean(Under20$FT), mean(Early20s$FT), mean(Late20s$FT), mean(Early30s$FT), mean(Above35$FT)) , "Free Throw Attempts" = c(mean(Under20$FTA), mean(Early20s$FTA), mean(Late20s$FTA), mean(Early30s$FTA), mean(Above35$FTA)))
#use gridExtra to combine the plots together

#Field Goals
FG <- qplot(Age, FG, data = stats) + ggtitle("Age vs Field Goals") + ylab("Field Goals")
FGA <- qplot(Age, FGA, data = stats) + ggtitle("Age vs Field Goal Attempts") + ylab("Field Goal Attempts")
FG. <- qplot(Age, FG., data = stats) + ggtitle("Age vs Field Goal Percentage") + ylab("Field Goal Percentage")
grid.arrange(tableGrob(FGDF),FG,FGA,FG., nrow = 2, widths= c(5,4))

  #Three Pointers
X3P <- qplot(Age, X3P, data = stats) + ggtitle("Age vs Three Pointers") + ylab("Three Pointers")
X3PA <- qplot(Age, X3PA, data = stats) + ggtitle("Age vs Three Point Attempts") + ylab("Three Point Attempts")
X3P. <- qplot(Age, X3P., data = stats) + ggtitle("Age vs Three Point Percentage") + ylab("Three Point Percentage")
grid.arrange(tableGrob(P3DF),X3P,X3PA,X3P., nrow = 2, widths= c(5,4))

#Two Pointers
X2P <- qplot(Age, X2P, data = stats) + ggtitle("Age vs Two Pointers") + ylab("Two Pointers")
X2PA <- qplot(Age, X2PA, data = stats) + ggtitle("Age vs Two Point Attempts") + ylab("Two Point Attempts")
X2P. <- qplot(Age, X2P., data = stats) + ggtitle("Age vs Two Point Percentage") + ylab("Two Point Percentage")
grid.arrange(tableGrob(P2DF),X2P,X2PA,X2P., nrow = 2, widths= c(5,4))

#Free Throws
FT <- qplot(Age, FT, data = stats) + ggtitle("Age vs Free Throws") + ylab("Free Throws")
FTA <- qplot(Age, FTA, data = stats) + ggtitle("Age vs Free Throw Attempts") + ylab("Free Throw Attempts")
FT. <- qplot(Age, FT., data = stats) + ggtitle("Age vs Free Throw Percentage") + ylab("Free Throw Percentage")
grid.arrange(tableGrob(FTDF),FT,FTA,FT., nrow = 2, widths= c(5,4))


```

# Team Reflection:

### Individual Reflections:

**David:**

At the start of this course my goal for 6 months after college was to be travelling in Europe, my 5 year goal was that I wanted to be analayzing statistics. These goals haven't changed for me, but I do feel that I am more commited to achieving my 5 year goal as a result of this course. I achieved at lease a basic understanding of R in this course, considering that this is my first experience with data science and coding, I'm pretty satisfied with this result. The advice that I would give myself at the beginning of this semester would be to get a good understanding of the examples in the book.

**Tanner:**

At the start of this course, my goal was for me to be able to develop the skills necessary to start doing my own projects and data science projects, in the hopes of finding a job in the future that maximizes my skills. I would say I achieved that goal. I have developed a solid understanding of the basics of R and data science, and am looking forward to trying my hand at some personal projects in the near future. My advice for myself would be to do more examples from the book or try developing a personal research project throughout the semester using the skills we were learning over time.

**Burke:**

At the start of this semester, my future was very uncertain. I had just switched out of engineering for psychology, and I know that was the right decision for me. However, I felt a strange pressure in this transition. I was nervous my family would be dissapointed in my new educational path. However, I have found a real love for psychology in the past few months, and I think that in five years I will find myself in graduate school. The reason I am taking statistics courses is because I've read that knowledge in statistics is necessary for when I have to do academic research later in my career. If I could go back, I would simply tell myself to put the extra time in for my classes whenever possible, as I know I could have done better this semester.

**Drew:**

My goal at the beginning of the year was to eventually get an internship or job doing sports statistics for a professional team. My experience in this course has only reinforced that goal for me. In our labs I enjoyed analyzing and plotting the sports data to draw conclusions and I believe it is something that I could see myself doing in the future. I was able to learn more about the R language and how to use it as well as other useful practices like tidy data and different forms of analysis. If I could give myself any advice at the beginning of the semester I would tell myself to get on the lab assignments before Thursdays. Other than that, I thoroughly enjoyed my experence in this class and believe it will have a significant effect on me in my future career.

**Brandon:**

At the start of this course my goal for 6 months was to learn if I actually like data science/stats and wonder if I wanted to pursue this path for my major. My 5 year goal is to be working for a US company in China. I am 90% sure I want to pursue data sci/stats as my major, althought my calculus grades are gonna pull my grades down a bit. I really enjoyed this course and it made my love for stats and data science grow siginificantly. The advice I would give myself at the beginning of this semester would be to google everything and explore R deeper than just what this class has to offer.